Model and results for identity signaling project

### 2020-05-22 paper/ms.tex
working through model description.

@@

\subsection{Signaling and receiving}

\subsection{Interaction and payoff}

\subsection{Social learning}

@@

Homophily not only determines

% The model is

% \begin{figure}[htpb]
% \begin{center}
% \caption{Model flow chart with relevant equations.}
% \label{fig:yo}
% \begin{tikzpicture}[]
%   \node[draw] at (0, 0) {}; 
% \end{tikzpicture}
% \end{center}
% \end{figure}

@@

@@

>>>>

\begin{table}[H]
  \begin{tabular}{ccp{5in}}
    Agent property & 
  \end{tabular}
\end{table}

@@

\subsection{Dynamics}

Each model iteration consist of three stages. First, agents signal their identities
to some fraction of all other agents. It is at this step that agent attitudes
$a_{ij}$ are set. Whether or not an agent receives a signal depends on the
efficiency of overt and covert signaling, $E_{ov}$ and $E_{cov}$, respectively.
\emph{Efficiency} is the proportion of agents in the population that receive
a given agent's signal depending on whether the agent is an overt or covert
signaler.  Signals are sent and received once every model iteration. 
If and agent does receive a signal

Then agents then
interact at most $N_I$ times with other agents. In each of the possible 
$N_I$ interactions, agents first match with one agent selected randomly from
the population. The probability these agents interact after matching depends
on the homophily, $w$, and the attitudes each agent has for the other. 
If a pair interacts, their payoff is determined by whether the agents are
similar, whether one dislikes the other, and the penalty for disliking,
$d$\footnote{(WILL INTRODUCE $\delta$ WITH A REFERENCE TO THE SI IN SUBSUBSEC
BELOW)}. The distinction between similarity and disliking is subtle, 
but necessary to include, as we explain in more detail below. After all
$N_I$ interaction rounds, each agent chooses a teacher to potentially learn
from. ``Learning'' is simply the adoption of either the signaling or
receiving strategy of the teacher by the learner---which strategy type is 
learned is chosen at random.

After several hundred iterations the prevalence of each signaling and
receiving strategy reaches an equilibrium (see SI dynamics plots).

\subsubsection{Signaling and receiving}

At the signaling and receiving step, agents form attitudes about one another if
they receive a signal from

\subsubsection{Interaction}

The two agents interact with different probabilities
depending on their attitudes about one another, and depending on the 
assumed homophily level. If  Note that as $w \rightarrow 0.5$,
agents never interact with agents they dislike and always interact with 
agents they like.

\begin{equation}
  \Pr(\text{$i$ interacts with $j$}) = \frac{1 + a_{ij} + a_{ji}}{2}
\end{equation}
\noindent

\subsubsection{Payoff}

Payoffs are determined based on the similarity bonus, $s$, and the disliking
penalty $d$\footnote{In the SI we consider the case where the disliking penalty 
when both agents dislike each other is different from $d$. We call this $\delta$
and let it vary from 0 to $2d$}.

\begin{table}[h]
  \centering
  \begin{tabular}{cccc}
    $a_{ij}$ & $a_{ji}$ & Similar? & Payoff $\pi_i = \pi_j$ \\
    \toprule
    1 & 1 & Yes & $1 + s$ \\
    1 & 0 & Yes & $1 + s$ \\
    1 & -1 & Yes & $1 + s - d$ \\
    1 & -1 & No & $1 + s - d$ \\
    0 & 0 & Yes & $1 + s$ \\
    0 & 0 & No & $1$ \\
    -1 & 0 & Yes & $1 + s - d$ \\
    -1 & 0 & No  & $1 - d$ \\
    -1 & -1 & Yes & $1 + s - 2d$ \\
    -1 & -1 & No  & $1 - 2d$ \\
  \end{tabular}
  \caption{Interaction payoff matrix. Certain conditions are impossible, and so are not
  shown. For instance, it is necessary for two agents to be similar for 
  one agent to like another agent (i.e.\ have $a_{ij} = 1$). Note that
  due to churlish receiving, two agents may be similar, but one dislikes the
  other while the other likes the one. There could be agents who dislike one
  another due to churlish receiving if they have never signaled each other.}
\end{table}

\subsubsection{Social learning}

All agents select another agent in the population to be a teacher. The 
learner agent probabilistically adopts its teacher's signaling or receiving 
strategy---which type of strategy might be learned is determined by a coin flip. 
The probability that the learner adopts the teacher's strategy is set 
calculated by the logistic sigmoid function applied to the quotient of 
teacher and learner payoffs accumulated in one signaling/receiving-interaction/payoff round.
Let $\pi_{i}$ be the accumulated payoff of agent $i$, and assume $i$ learns
from $j$. Set $p_{ij} = \pi_{j} / \pi_i$. 
The probability the learner will adopt the teacher's strategy is

\begin{equation}
  \Pr(\text{$i$ learns from $j$}) = \frac{1}{1 + e^{-\beta(p_{ij} - \alpha)}}.
\end{equation}
\noindent
$\alpha$ and $\beta$ are parameters shift and scale the sigmoid curve, respectively. 
$\alpha$ is the size of $p_{ij}$ such that $\Pr(\text{$i$ learns from $j$}) = 0.5$. % $P(\text{$i$ learns from $j$}) = 0.5$.
For example, if we set $\alpha = 2$, then the teacher's accumulated payoff
must be twice that of the learner's for a 50\% chance that the learner 
adopts the teacher's strategy. 
For the main results we set $\alpha=1.25$. $\beta$ sets how quickly 
$\Pr(\text{$i$ learns from $j$})$ increases overall. Larger $\beta$ means that
if $p_{ij} < 0.5$ then $\Pr(\text{$i$ learns from $j$})$ is closer to 0 than smaller $\beta$. 
$\Pr(\text{$i$ learns from $j$})$ is closer to 1 for larger than for smaller $\beta$. For the main
results we set $\beta = 12$. These settings are rather arbitrary, (SO WE NEED
TO PROVIDE SENSITIVITY CHECKS IN THE SUPPLEMENT).

@@

\subsection{Outcome measures}

We are primarily interested in the prevalence of covert signaling at time
$t$. Usually, $t=T$, where $T$ is the number of iterations chosen for our 
computational experiments. The prevalence of overt signalers is the complement 
of the prevalence of covert signalers. One important result shows that the
prevalence of covert signalers at $t=T$ is positively correlated with the
prevalence of generous receivers at $t=T$.

\begin{table}[H]
  \centering
  \begin{tabular}{rp{3in}} %p{5in}}
    Symbol & Measure \\ %& Description \\
    \toprule
      $\rho_{cov}$ & Prevalence of covert signalers\\
    $1 - \rho_{cov} = \rho_{ov}$ & Prevalence of overt signalers  \\
     $\rho_{gen}$ & Prevalence of generous receivers \\
    $1 - \rho_{gen} = \rho_{ch}$ & Prevalence of churlish receivers \\
    $\rho_i^{minor}$, $\rho_i^{maj}$ & Prevalence of strategy $i \in \{cov, ov, gen, ch\}$ amongst minority/majority \\
    $\Pr(\text{$i$ invades})$ & Probability strategy $i$ invades 
  \end{tabular}
  \caption{Summary of outcome measures. To indicate any of these values at
  a particular time, we add an additional subscript. For example, $\rho_{cov,t}^{maj}$
  is the prevalence of covert signalers among the majority at time $t$.}
  \label{tab:outcomeMeasures}
\end{table}

@@

Theoretically, conditions favorable for covert
signalers should also be favorable for generous receivers (TRUE??). Furthermore, 
an increase in covert signalers increases the pressure towards generous receiving.
With many covert signalers it is likely an agent will get no signal on a 
turn because they were receiving a signal from a covert signaler. Disliking
incurs a penalty, so it is better to remain neutral.

Our computational results support this theoretical claim. As the heatmaps
below reveal... In the SI we show regressions of $\rho_{cov,T} \sim \rho_{gen,T}$.

@@

\section{Model implementation}

In order to aid evaluation of our model code, we provide notes below on 
our model implementation. The code is freely available on GitHub at
\url{https://github.com/mt-digital/identity-signaling}.
